---
title: Examples of Expanding the Orbis Network out from Rome
output:
  html_document
---

## Introduction

A few examples of calculating graph distance from Rome according to http://orbis.stanford.edu . Data from http://purl.stanford.edu/mn425tz9757 .

Adapted from code and comments developed in [NYU/ISAW's](http://isaw.nyu.edu) "Mapping and Data  Visualization" Spring 2015 seminar.


## Load some libraries

```{r libraries, message=FALSE, warning=FALSE}
#load libraries
library(curl)
library(igraph)
```

## Load data and make graph
```{r data, message = FALSE, warning=FALSE}
# cawd is an R package of "Collected Ancient World Data"
# http://github.com/sfsheath/cawd
library(cawd) 

# igraph like a 'label' column
orbis.nodes$label <- orbis.nodes$title

#setting the network
o.nw  <- graph.data.frame(orbis.edges, vertices=orbis.nodes, directed=TRUE)

# this now plot()able but it's so big that it's too messy to be useful
```

## Calculate shortest paths to Rome
```{r calculate, message=FALSE, warning=FALSE}
#finding the shortest paths referring to Roma, by expenses, and assigning to a new object
shortest.paths(o.nw,V(o.nw)[V(o.nw)$title == "Roma"], weights = E(o.nw)$expense) -> o.nw.torome


o.nw <- set.vertex.attribute(o.nw, "torome", index = V(o.nw), value = o.nw.torome)
```

## Divide into equal fiftieths (or choose your number)
```{r quartiles}
# first print it
quantile(V(o.nw)$torome, probs = 1:50/50)

# now keep it for future use
quantile(V(o.nw)$torome, probs = 1:50/50) -> breaks

```

## Now subgraph by various values of torome
This could be turned into a trivial function but I just copy-and-paste for clarity. For now...

In all of the following examples, the V() function, which lists verteces in an igraph graph, is used to select verteces accoring to their "distance" from Rome accoring to the Orbis graph.

Note that I am just increasing the index into 'breaks[]'.

### Very, very "close"
```{r one, message=FALSE, warning=FALSE}

tmp.nw <- subgraph(o.nw, v = V(o.nw)[torome <= breaks[1]])

plot(tmp.nw,
     edge.arrow.size = 0,
     vertex.label.cex = .6
     )
```

### Very "close"
```{r two, message=FALSE, warning=FALSE}

tmp.nw <- subgraph(o.nw, v = V(o.nw)[torome <= breaks[2]])

plot(tmp.nw,
     edge.arrow.size = 0,
     vertex.label.cex = .6
     )
```

### "close", now with label color = 'red'
```{r three, message=FALSE, warning=FALSE}

tmp.nw <- subgraph(o.nw, v = V(o.nw)[torome <= breaks[3]])

plot(tmp.nw,
     edge.arrow.size = 0,
     vertex.label.cex = .6,
     vertex.label.color = 'red'
     )
```

### Getting messy so smaller vertex.size...
```{r four, message=FALSE, warning=FALSE}

tmp.nw <- subgraph(o.nw, v = V(o.nw)[torome <= breaks[5]])

plot(tmp.nw,
     edge.arrow.size = 0,
     vertex.label.cex = .6,
     vertex.size = 2
     )
```

### Last one, the closest 20%...
```{r five, message=FALSE, warning=FALSE}

tmp.nw <- subgraph(o.nw, v = V(o.nw)[torome <= breaks[10]])

plot(tmp.nw,
     edge.arrow.size = 0,
     vertex.label.cex = .6,
     vertex.size = 2
     )
```


## Next up, turning these into maps...

